---
title: "Stratified avgs"
author: "Abby Lewis"
date: "2022-11-22"
output: html_document
---

This file loads temp/DO/productivity data, calculates summary statistics for the dataset, and calculates mean concentrations of each focal variable during the summer stratified period. All code chunks must be run in order.

Table of contents:
Step 1: Load data and packages
Step 2: Summary statistics
Step 3: Filter to stratified period
Step 4: Calculate metalimetic boundaries and average values within each layer
Step 5: Add buoyancy frequency, hypolimnetic SA:vol ratio, and chl-a during the stratified period
Step 6: Plot stratification duration
Step 7: Plot data availability


Step 1: Load data and packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rLakeAnalyzer)
source("thermo.depth.density.R")

## Load productivity data and metadata from EDI
p <- read.csv("https://pasta-s.lternet.edu/package/data/eml/edi/1029/9/0ece9d7b67cd49741ed7ee60192832e4")
lat_long <- read.csv("https://pasta-s.lternet.edu/package/data/eml/edi/1029/9/fadd3eaa25b5fdd1fc4efba70e660579")
#Load interpolated DO data file, created by "02 - Temp and DO interpolation.Rmd"
do <- read.csv("../Compiled data/temp_o2_interpolated.csv")
```


Step 2: Summary statistics
```{r}
#TN and DOC summary stats
doc <- p%>%
  filter(!is.na(DOC_mgL))%>%
  dplyr::summarise(n = length(unique(LakeID)))
doc[1,1]

tn <- p%>%
  filter(!is.na(TN_ugL))%>%
  dplyr::summarise(n = length(unique(LakeID)))
tn[1,1]

#Merge P, DO, and lake info (from lat long file)
full <- p%>%
  full_join(do, by = c("Date","LakeID", "Depth_m"))%>%
  full_join(lat_long, by = c("LakeID"))%>%
  mutate(Date = as.Date(Date))%>%
  filter(!is.na(MaximumDepth_m)&MaximumDepth_m>=6.4)
```


Step 3: Filter to stratified period
```{r}
#Create a Date_22 column, so that I can filter to only the end of summer
full$Date_22 <- full$Date
year(full$Date_22) <-2022

#There are a few rows with missing dates because these lakes are in the lat long database and not do/p
full%>%
  filter(is.na(as.Date(Date)))
#Remove these NAs
full <- full%>%
  filter(!is.na(Date),
         !is.na(Latitude_DD))

###
#Determine start and end of stratification, then thermo depths
###

full_format <- full%>%
  mutate(Date_unif = as.Date(ifelse(Latitude_DD>=0,Date,Date+months(6)),origin = "1970-01-01"))%>%
  group_by(Date, Date_unif, LakeID, Depth_m)%>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T))%>%
  filter(!is.na(Temp_C))%>%
  ungroup()%>%
  group_by(Date, Date_unif, LakeID)%>%
  dplyr::summarize(epi_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[1],
            hypo_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[2],
            max_depth = max(Depth_m),
            thermo = thermo.depth.density(Temp_C,Depth_m, mixed.cutoff = 0.1, seasonal = F))%>% #custom thermocline function
  mutate(Year= year(Date_unif),
         unstrat = as.numeric(is.na(thermo)))
#Start of stratification
date_start <- full_format%>%
  group_by(Year,LakeID)%>%
  filter(unstrat==1,
         month(Date_unif)<=7)%>%
  filter(Date==max(Date_unif))%>%
  dplyr::select(Date, Year, LakeID)%>%
  rename(Date_start = Date)
#End of stratification
date_end <- full_format%>%
  group_by(Year,LakeID)%>%
  filter(unstrat==1,
         month(Date_unif)>7)%>%
  filter(Date==min(Date_unif))%>%
  dplyr::select(Date, Year, LakeID)%>%
  rename(Date_end = Date)
#Stratification duration
strat_dur <- date_start%>%
  full_join(date_end)

#Filter to stratified period
full_trimmed <- full%>%
  mutate(Date_unif = as.Date(ifelse(Latitude_DD>=0,Date,Date+months(6)),origin = "1970-01-01"),
         Year = year(Date_unif))%>%
  full_join(strat_dur)%>%
  filter(is.na(Date_start)|Date_unif>Date_start,
         is.na(Date_end)|Date_unif<Date_end,
         month(Date_unif)>3,
         month(Date_unif)<11)
```


Step 4: Calculate metalimetic boundaries and average values within each layer
```{r}
#Thermocline depths during stratified period
thermo_depths <- full_format%>%
  full_join(strat_dur)%>%
  filter(is.na(Date_start)|Date_unif>Date_start,
         is.na(Date_end)|Date_unif<Date_end)%>%
  group_by(Year,LakeID)%>%
  dplyr::summarize(epi_depth = mean(epi_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T),
            count_unstrat = sum(unstrat),
            n= n())

thermo_depths_sum <- thermo_depths%>%
  group_by(LakeID)%>%
  dplyr::mutate(count_unstrat_tot = sum(count_unstrat),
                n = sum(n))%>%
  filter((count_unstrat_tot/n) <0.1,
         count_unstrat == 0)%>%
  group_by(LakeID,Year)%>%
  dplyr::summarize(epi_sd = sd(epi_depth, na.rm = T),
            epi_depth = mean(epi_depth, na.rm = T),
            hypo_sd = sd(hypo_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T))

full_with_thermo <- full_trimmed%>%
  full_join(thermo_depths_sum)%>%
  filter(!is.na(epi_depth),
         !is.na(hypo_depth)
         )
write.csv(full_with_thermo,"../Compiled data/Stratified_period_data_with_thermo.csv",row.names = F)

summer_layers <- full_with_thermo%>%
  mutate(Layer = ifelse(!is.na(Depth_m)&Depth_m<epi_depth, "EPI", NA),
         Layer = ifelse(is.na(Depth_m)&!is.na(Interval)&Interval=="EPILIMNION","EPI",Layer),
         Layer = ifelse(!is.na(Depth_m)&Depth_m>hypo_depth,"HYPO",Layer),
         Layer = ifelse(is.na(Depth_m)&!is.na(Interval)&Interval=="HYPOLIMNION","HYPO",Layer),
         Layer = ifelse(!is.na(Depth_m)&Depth_m<hypo_depth&Depth_m>epi_depth, "META",Layer),
         Layer = ifelse(is.na(Depth_m)&!is.na(Interval)&Interval=="METALIMNION","META",Layer))%>%
  filter(!is.na(Layer))

summer_avgs <- summer_layers%>%
  group_by(LakeID,Year, Layer)%>%
  mutate(start_date = as.Date(paste0(Year,"-08-01")))%>%
  dplyr::summarize(TP_ugL = mean(TP_ugL, na.rm = T),
            TP_date = as.numeric(mean(Date[!is.na(TP_ugL)],na.rm=T)-unique(start_date)),
            DOC_mgL = mean(DOC_mgL, na.rm = T),
            DOC_date = as.numeric(mean(Date[!is.na(DOC_mgL)],na.rm=T)-unique(start_date)),
            DO_mgL = mean(DO_mgL, na.rm = T),
            DO_date = as.numeric(mean(Date[!is.na(DO_mgL)],na.rm=T)-unique(start_date)),
            Chla_ugL = mean(Chla_ugL, na.rm = T),
            Chla_date = as.numeric(mean(Date[!is.na(DO_mgL)],na.rm=T)-unique(start_date)),
            Temp_C = mean(Temp_C, na.rm = T),
            Temp_date = as.numeric(mean(Date[!is.na(Temp_C)],na.rm=T)-unique(start_date)),
            TN_ugL = mean(TN_ugL, na.rm = T),
            TN_date = as.numeric(mean(Date[!is.na(TN_ugL)],na.rm=T)-unique(start_date)),
            hypo_depth = unique(hypo_depth))
```


Step 5: Add buoyancy frequency, hypolimnetic SA:vol ratio, and chl-a during the stratified period
```{r}
###
#Calculate buoyancy frequency
###

buoyancy <- full_trimmed%>%
  group_by(Date, LakeID, Depth_m)%>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T))%>%
  filter(!is.na(Temp_C))%>%
  ungroup()%>%
  group_by(Date, LakeID)%>%
  dplyr::summarize(buoyancy_freq = max(buoyancy.freq(Temp_C,Depth_m),na.rm=T))%>%
  mutate(Year=year(Date))%>%
  group_by(Year,LakeID)%>%
  dplyr::summarize(buoyancy_freq = mean(buoyancy_freq,na.rm=T))
#Add to df
summer_avgs <- summer_avgs%>%
  left_join(buoyancy)

###
#Calculate surface area:volume ratio of the hypolimnion
###

#Load bathymetry
lake_bats.raw <- read.csv("https://pasta-s.lternet.edu/package/data/eml/edi/1029/9/ba733454c29a5f026efed05a6e6ef75b")%>%
  filter(Area_m2>0)%>%
  group_by(LakeID)%>%
  mutate(methods_n = length(unique(Method)))%>%
  filter(methods_n==1|Method=="model")%>%
  ungroup()
#Calculate ratio
ratio <- thermo_depths_sum%>%
  mutate(Depth_m = round(hypo_depth))%>%
  dplyr::select(LakeID,Year,Depth_m)%>%
  left_join(lake_bats.raw, by = c("LakeID", "Depth_m"))%>%
  mutate(SA_vol_ratio = Area_m2/CumulativeVolume_m3)%>%
  dplyr::select(LakeID,Year,SA_vol_ratio)
#Add to df
summer_avgs <- summer_avgs%>%
  left_join(ratio)

###
#Load published chl-a data
###

chl <- read.csv("../External data/ChlaData.csv") #https://knb.ecoinformatics.org/view/doi:10.5063/F1JH3JKZ
lacking_data <- summer_avgs%>%
  group_by(LakeID)%>%
  summarize(n = sum(is.na(Chla_ugL)),
            tot = n())%>%
  filter(n==tot)
name_harmonizer <- read.csv("chla_harmonizer.csv") #Manually harmonized lake names

harm_short <- name_harmonizer%>%
  filter(ID%in%lacking_data$LakeID)%>%
  filter(nchar(UniqueLakeName)>1)

harm_short%>%filter(lat<0) #don't have to worry about southern hemisphere

published_chla <- chl%>%
  mutate(Year = as.numeric(Year))%>%
  filter(UniqueLakeName%in%harm_short$UniqueLakeName)%>%
  filter(!is.na(Month),
         (Month%in%5:8&Lat>0)|
           Month%in%c(11,12,1,2)&Lat<0)%>%
  left_join(harm_short%>%dplyr::select(ID,UniqueLakeName))%>%
  rename(LakeID=ID)%>%
  group_by(LakeID,Year)%>%
  dplyr::summarise(Chla_ugL = mean(ChlValues,na.rm = T)*1000)%>%
  mutate(Layer="EPI")

published_chla <- published_chla%>%
  rename(Chla_ugL_published = Chla_ugL)

#If we do not currently have any Chla data, add published data
summer_avgs <- summer_avgs%>%
  full_join(published_chla)%>%
  mutate(Chla_ugL = ifelse(is.na(Chla_ugL),Chla_ugL_published,Chla_ugL))%>%
  dplyr::select(-Chla_ugL_published)

#All done!!
write.csv(summer_avgs,"../Compiled data/stratified_averages.csv")
```


Step 6: Plot stratification duration
```{r}
jpeg("../Figures/Stratification.jpg", width = 6, height = 4, res = 300, units = "in")
strat_dur%>%
  left_join(lat_long, by = c("LakeID"="LakeID"))%>%
  group_by(LakeID)%>%
  dplyr::summarize(min_date = mean(yday(as.Date(Date_start, origin = "1970-01-01")), na.rm=T),
            max_date = mean(yday(as.Date(Date_end, origin = "1970-01-01")), na.rm=T))%>%
  pivot_longer(cols = c(min_date,max_date))%>%
  group_by(name)%>%
  mutate(median = median(value,na.rm=T),
         median_hjust = ifelse(name == "min_date",-.1,1.1))%>%
  mutate(name = factor(name, levels = c("min_date","max_date"), labels = c("Latest non-stratified profile\n(spring)","Earliest non-stratified profile\n(fall)")))%>%
  ggplot(aes(x = value, fill = name))+
  geom_density()+
  geom_vline(aes(xintercept = median),lty = "dashed")+
  annotate("rect",ymin = 0,ymax=Inf,xmin = yday(as.Date("2022-07-15")),xmax = yday(as.Date("2022-08-31")),alpha=0.2,fill="black")+
  geom_vline(aes(xintercept = yday(as.Date("2022-3-01"))))+
  geom_vline(aes(xintercept = yday(as.Date("2022-12-01"))))+
  geom_text(aes(x = median, y = 0.019, hjust = median_hjust,label = paste0("median = ",round(median))), size = 3)+
  scale_fill_manual(values = c("pink","chocolate4"), name = "")+
  xlab("Day of year (Northern Hemisphere)\nSouthern Hemisphere dates are offset by 6 months")+
  ylab("Density")+
  theme_bw()+
  theme(legend.position = "bottom")
dev.off()
```


Step 7: Plot data availability
```{r}
data_for_plot <- full%>%
  dplyr::select(Date,Depth_m,Temp_C, DO_mgL,LakeID, TP_ugL,Chla_ugL)%>%
  pivot_longer(cols = c(Temp_C, DO_mgL,TP_ugL,Chla_ugL))%>%
  group_by(LakeID,name)%>%
  dplyr::summarize(n_year = length(unique(year(Date)[!is.na(value)])))%>%
  group_by(name)%>%
  filter(!n_year==0)%>%
  mutate(ten_plus = as.numeric(n_year>=10))%>%
  dplyr::summarize(tot_ten = sum(ten_plus),
            tot = n())%>%
  pivot_longer(cols = c(tot,tot_ten), names_to = "var")

layered_summary <- summer_avgs%>%
  dplyr::select(Temp_C, DO_mgL,LakeID, TP_ugL,Chla_ugL,Layer,Year)%>%
  filter(!Layer == "META")%>%
  pivot_longer(cols = c(Temp_C, DO_mgL,TP_ugL,Chla_ugL))%>%
  group_by(LakeID,name,Layer)%>%
  dplyr::summarize(n_year = length(unique(Year[!is.na(value)])))%>%
  group_by(name, Layer)%>%
  filter(!n_year==0)%>%
  mutate(ten_plus = as.numeric(n_year>=10))%>%
  dplyr::summarize(tot_ten = sum(ten_plus),
            tot = n())%>%
  pivot_longer(cols = c(tot,tot_ten), names_to = "var")

library(viridis)
jpeg("../Figures/data_availability_sum.jpg",width = 6, height = 2.5, units = "in", res = 300)
all_data_plot <- data_for_plot%>%
  mutate(Layer = "All data")%>%
  full_join(layered_summary)%>%
  mutate(var = ifelse(var=="tot","Any data","≥10 years"),
         var = factor(var, levels = c("Any data","≥10 years")),
         name = ifelse(name=="Chla_ugL","Chl-a (µg/L)",
                       ifelse(name=="DO_mgL","DO (mg/L)",
                              ifelse(name=="Temp_C","Temperature (ºC)",
                                     ifelse(name=="TP_ugL","TP (µg/L)",NA)))),
         name = factor(name, levels = c("Temperature (ºC)","DO (mg/L)","TP (µg/L)","Chl-a (µg/L)")),
         Layer = ifelse(Layer=="EPI",
                        "Epilimnion",
                        ifelse(Layer=="HYPO","Hypolimnion",Layer)),
         Layer = factor(Layer))%>%
  #filter(!name=="Chl-a (µg/L)"|!Layer=="Late-summer hypo.")%>%
  ggplot(aes(x = var,y=value, fill = Layer))+
  geom_bar(stat= "identity", position = "dodge", color = "black",lwd=.3)+
  facet_grid(rows = ~name, space = "free_x")+
  theme_bw()+
  scale_fill_manual(values = c("grey25","#aee2fa","#5ca4d9"))+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom")+
  ylab("Number of lakes")
all_data_plot
dev.off()
```