---
title: "ASL DO Demand calcs"
author: "Abby Lewis"
date: "2022-08-29"
output: html_document
---

This file calculates volume-weighted hypolimnetic oxygen demand during the stratified period at each lake. 

Table of contents:
Step 1: Load data and packages
Step 2: Calculate the rate of change in volume-weighted oxygen concentrations during the stratified period at each lake (VHOD)
Step 3: Calculate temperature-corrected oxygen demand values
Step 4: Calculate trends in oxygen demand over time


Step 1: Load data and packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rLakeAnalyzer)
library(openair)

#Load data and stratification boundaries. This file is created by "03 - Stratified avgs.Rmd"
full_with_thermo <- read.csv("../Compiled data/Stratified_period_data_with_thermo.csv")%>%
  mutate(Date=as.Date(Date))
#Load lake metadata
lat_long <- read.csv("https://pasta-s.lternet.edu/package/data/eml/edi/1029/9/fadd3eaa25b5fdd1fc4efba70e660579")
```


Step 2: Calculate the rate of change in volume-weighted oxygen concentrations during the stratified period at each lake (VHOD)
```{r}
#prepare bathymetry
#lake with only one method (data/model)
lake_bats <- read.csv("https://pasta-s.lternet.edu/package/data/eml/edi/1029/9/ba733454c29a5f026efed05a6e6ef75b")%>%
  filter(Area_m2>0)%>%
  group_by(LakeID)%>%
  mutate(methods_n = length(unique(Method)))%>%
  filter(methods_n==1|Method=="model")%>%
  ungroup()

#Check bathymetric data availability
unique(lake_bats$LakeID)[!unique(lake_bats$LakeID)%in%unique(lat_long$LakeID)]
unique(lat_long$LakeID)[!unique(lat_long$LakeID)%in%unique(lake_bats$LakeID)] #missing full bathymetry for 37 lakes

#Total n to start: 609
length(unique(full_with_thermo$LakeID))

#Calculate volume-weighted oxygen concentrations
vw_do <- full_with_thermo%>%
  full_join(lake_bats, by = c("LakeID","Depth_m"))%>%
  filter(!is.na(IntervalVolume_m3))%>%
  filter(Depth_m>hypo_depth)%>%
  mutate(DO_mass = DO_mgL*IntervalVolume_m3,
         Temp_total = Temp_C*IntervalVolume_m3,
         Year = year(Date))%>%
  group_by(Date, LakeID)%>%
  dplyr::summarize(DO_tot = sum(DO_mass),#sum across all hypolimnetic depths
                   vol_tot = sum(IntervalVolume_m3),
                   DO_mgL_vol = DO_tot/vol_tot,#Divide by hypolimnetic volume
                   DO_mgL_area = DO_tot/max(Area_m2),
                   Temp_C = sum(Temp_total)/vol_tot)%>%
  mutate(Year= year(Date))%>%
  group_by(LakeID, Year)%>%
  filter(!is.na(DO_mgL_vol))%>%
  arrange(LakeID, Date)%>%
  mutate(low_point = ifelse(sum(DO_mgL_vol<1)==0,#Identify the lowest DO value for a given year
                            Date[which.min(DO_mgL_vol)],
                            first(Date[DO_mgL_vol<1])))%>%
  filter(is.na(low_point)|Date<=low_point)%>% #Remove days after the lowest DO value
  ungroup()

length(unique(vw_do$LakeID)) #581 lakes left

#Calculate the rate of change in volume-weighted concentrations using lm
vw_do_demand <- vw_do%>%
  mutate(Year = year(Date))%>%
  group_by(Year, LakeID)%>%
  dplyr::summarize(n = n(),
            DO_demand_mgLd = -lm(DO_mgL_vol~Date)$coefficients[2],
            AHOD_mgLd = -lm(DO_mgL_area~Date)$coefficients[2],
            p_vol = summary(lm(DO_mgL_vol~Date))$coefficients[8],
            p_area = summary(lm(DO_mgL_area~Date))$coefficients[8],
            r2_vol = summary(lm(DO_mgL_vol~Date))$r.squared,
            r2_area = summary(lm(DO_mgL_area~Date))$r.squared,
            Temp_C = mean(Temp_C, na.rm = T))%>%
  filter(n>=3)#Need at least 3 points
length(unique(vw_do_demand$LakeID[!is.na(vw_do_demand$DO_demand_mgLd)])) #418

#How many rows get removed by filtering so R2>0.5? 246
removed_by_r2 <- vw_do_demand%>%
  filter(r2_vol>.5)
nrow(removed_by_r2[!is.na(removed_by_r2$DO_demand_mgLd),]) #5139
nrow(vw_do_demand[!is.na(vw_do_demand$DO_demand_mgLd),]) #5385
5385-5139
(5385-5139)/5385

#How many rows get removed by filtering so oxygen demand is positive? 0
removed_by_pos <- removed_by_r2%>%
  filter(DO_demand_mgLd>0)
nrow(removed_by_pos[!is.na(removed_by_pos$DO_demand_mgLd),])

#Finish QAQC
good_hod <- vw_do_demand%>%
  group_by(LakeID)%>%
  filter(!is.na(DO_demand_mgLd),
         r2_vol>0.5
         )%>%
  dplyr::summarize(n_neg = sum(DO_demand_mgLd<0),
                   n = n(),
                   pct = n_neg/n*100)%>%
  filter(pct<=10)
length(unique(good_hod$LakeID)) #414

vw_do_demand_qaqc <- removed_by_pos%>%
  filter(LakeID %in% good_hod$LakeID)

nrow(vw_do_demand_qaqc[!is.na(vw_do_demand_qaqc$DO_demand_mgLd),]) #5139
length(unique(vw_do_demand_qaqc$LakeID[!is.na(vw_do_demand_qaqc$DO_demand_mgLd)])) #414

#Success!
write.csv(vw_do_demand_qaqc, "../Compiled data/VW oxygen demand points.csv", row.names = F)
```


Step 3: Calculate temperature-corrected oxygen demand values
```{r}
vhod5 <- vw_do_demand_qaqc%>%
  mutate(Temp_factor = ((Temp_C)/10)^0.65,
         VHOD5_mgLd = DO_demand_mgLd/Temp_factor)%>%
  filter(!is.na(VHOD5_mgLd))%>%
  dplyr::select(Year,LakeID,Temp_factor,VHOD5_mgLd)

write.csv(vhod5, "../Compiled data/VHOD5 points.csv", row.names = F)
```


Step 4: Calculate trends in oxygen demand over time. This is not currently part of the manuscript, but I thought the analysis was interesting
```{r}
#At one point, I was interested in whether trends are different before and after a given year, 
#so the function below is set up to accommodate calculating trends on either side of a given year
#not currently using, and may not work well at this point
split_year = 1800

#Function
sen_slope_custom <- function(df,var){
  output = df%>%
    mutate(era = ifelse(Year<split_year,"before","after"))%>%
    group_by(LakeID,era)%>%
    dplyr::summarize(n = n(),
              trend = NA,
              sig = NA,
              min_year = NA,
              max_year = NA)
  for(lake in unique(df$LakeID)){
    for(era in c("before","after")){
      print(paste(lake,era))
      filt = df%>%
        filter(LakeID==lake)%>%
        mutate(date = as.POSIXct(paste0(Year,"-01-01")))
      if(era == "before"){
        filt = filt%>%
          ungroup()%>%
          filter(Year<split_year)%>%
          dplyr::select(-Year)
      }else{
        filt = filt%>%
          ungroup()%>%
          filter(Year>split_year)%>%
          dplyr::select(-Year)
      }
      if(length(unique(year(filt$date)))>=10){#Only calculate a trend if there are 10 years of data
        sen = TheilSen(filt, pollutant = var, avg.time = "year", plot = F)$data$res2
        output$trend[output$LakeID==lake&output$era == era]<-sen$slope[1]
        output$sig[output$LakeID==lake&output$era == era]<-sen$p[1]
        output$min_year[output$LakeID==lake&output$era == era]<-min(year(filt$date))
        output$max_year[output$LakeID==lake&output$era == era]<-max(year(filt$date))
      }
    }
  }
  return(output)
}

#Calculate trends
vw_demand_sen <- sen_slope_custom(vw_do_demand_qaqc,"DO_demand_mgLd")
vhod5_sen <- sen_slope_custom(vhod5,"VHOD5_mgLd")

#Plot
vw_demand_sen%>%
  filter(!is.na(trend),
         sig<0.05)%>%
  mutate(type = ifelse(trend<0,"Decreasing","Increasing"))%>%
  ggplot(aes(x = type))+
  geom_bar()+
  xlab("")+
  ggtitle("Of lakes with significant demand trends:")

vw_demand_sen%>%
  ungroup()%>%
  filter(!is.na(trend))%>%
  mutate(n_tot = n())%>%
  filter(sig<0.05)%>%
  mutate(type = ifelse(trend<0,"Decreasing","Increasing"))%>%
  group_by(type)%>%
  dplyr::summarize(n = n(),
                   pct = n/unique(n_tot)*100)

vhod5_sen%>%
  filter(!is.na(trend),
         sig<0.05)%>%
  mutate(type = ifelse(trend<0,"Decreasing","Increasing"))%>%
  ggplot(aes(x = type))+
  geom_bar()+
  xlab("")+
  ggtitle("Of lakes with significant temp-corrected demand trends:")

vhod5_sen%>%
  ungroup()%>%
  filter(!is.na(trend))%>%
  mutate(n_tot = n())%>%
  filter(sig<0.05)%>%
  mutate(type = ifelse(trend<0,"Decreasing","Increasing"))%>%
  group_by(type)%>%
  dplyr::summarize(n = n(),
                   pct = n/unique(n_tot)*100)

vhod5_sen%>%
  filter(!is.na(trend),
         sig<0.05)%>%
  rename(trend_vhod5 = trend)%>%
  dplyr::select(-sig,-n,-min_year,-max_year)%>%
  full_join(vw_demand_sen%>%
              filter(!is.na(trend),
                     sig<0.05)%>%
              rename(trend_vhod = trend)%>%
              dplyr::select(-sig,-n,-min_year,-max_year))

vhod5_sen%>%
  rename(trend_vhod5 = trend)%>%
  dplyr::select(-sig,-n,-min_year,-max_year)%>%
  full_join(vw_demand_sen%>%
              rename(trend_vhod = trend)%>%
              dplyr::select(-sig,-n,-min_year,-max_year))%>%
  ggplot(aes(x = trend_vhod, y = trend_vhod5))+
  geom_point()+
  geom_abline(slope=1)

median <- median(vhod5_sen$trend, na.rm = T)
vhod5_sen%>%
  ggplot(aes(x = trend))+
  geom_density()+
  geom_vline(xintercept = median, lty = "dashed")+
  geom_vline(xintercept = 0)

vhod5_sen_format <- vhod5_sen%>%
  mutate(var = "VHOD5",
         Layer = "HYPO",
         season = "Summer")

vw_demand_sen_format <- vw_demand_sen%>%
  mutate(var = "DO_demand",
         Layer = "HYPO",
         season = "Summer")

write.csv(vhod5_sen_format, "../Compiled data/VHOD5 trends.csv", row.names = F)
write.csv(vw_demand_sen_format, "../Compiled data/VW oxygen demand trends.csv", row.names = F)
```
