---
title: "Air temp analysis"
author: "Abby Lewis"
date: "2023-05-22"
output: html_document
---

This file analyzes the effect of variation in air temperature on hypolimnetic temperature and DO. 

Table of contents:
Step 1: Load packages and data
Step 2: Generate temp anomaly summary fig
Step 3: Analyze the effect of monthly air temperature on hypolimnetic DO, VHOD, and temp dynamics


Step 1: Load packages and data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(mosaic)
library(ggallin)
library(ppcor)
library(multcompView)
library(ggpubr)
library(viridis)
library(MASS)

#Load saved data
with_temp = read.csv("../Compiled data/All_data_annual.csv")
```


Step 2: Generate temp anomaly summary fig
```{r}
zscore <- function(data, na.rm = T){(data-mean(data, na.rm = na.rm))/sd(data, na.rm = na.rm)}
do_anomaly_stat <- with_temp%>%
  group_by(LakeID)%>%
  mutate(hypo_do_anom = DO_mgL_HYPO-mean(DO_mgL_HYPO, na.rm = T),
         spr_temp = (temp_mar+temp_apr)/2,
         spr_temp_anom = spr_temp - mean(spr_temp, na.rm = T))

cor_data <- do_anomaly_stat%>%
  dplyr::select(hypo_do_anom,spr_temp_anom,Year)%>%
  na.omit()

pcor.test(cor_data$hypo_do_anom, cor_data$spr_temp_anom,cor_data$Year,method = "spearman")$estimate

jpeg("../Figures/DO and air temp anomaly - spring.jpg", width = 5, height = 3, res = 300, units = "in")
do_air_sum <- do_anomaly_stat%>%
  filter(!is.na(spr_temp_anom),
         !is.na(hypo_do_anom))%>%
  group_by(LakeID)%>%
  mutate(hypoxic = ifelse(max(DO_mgL_HYPO,na.rm = T)<1,"y","n"))%>%
  ungroup()%>%
  filter(hypoxic=="n")%>%
  mutate(temp_anom_sum = ifelse(spr_temp_anom>0.6,"> 0.6",
                                ifelse(spr_temp_anom<(-0.6),"< -0.6","between -0.6 \nand 0.6")),
         temp_anom_sum = factor(temp_anom_sum, levels = c("< -0.6", "between -0.6 \nand 0.6", "> 0.6")),
         do_anom_sum = ifelse(hypo_do_anom>0.2,"> 0.2",
                                ifelse(hypo_do_anom<(-0.2),"< -0.2","between -0.2 \nand 0.2")),
         do_anom_sum = factor(do_anom_sum, levels = c("> 0.2", "between -0.2 \nand 0.2", "< -0.2")))%>%
  group_by(temp_anom_sum)%>%
  mutate(n=n())
length(unique(do_air_sum$LakeID))
do_air_sum%>%
  group_by(temp_anom_sum)%>%
  mutate(temp_n = n())%>%
  group_by(temp_anom_sum,do_anom_sum)%>%
  dplyr::summarize(do_n = n(),
            pct = do_n/unique(temp_n)*100)
do_air_sum%>%
  ggplot(aes(fill = do_anom_sum, x = temp_anom_sum))+
  geom_bar(position = "fill")+
  geom_hline(yintercept = .5, color = "grey80",lty = "dashed")+
  geom_text(aes(y = .9, label = paste0("n = ",n)), color = "grey50")+
  labs(fill="Hypolimnetic \nDO anomaly (mg/L)")+
  xlab("Spring air temperature anomaly (ºC)")+
  ylab("Percent of lake-years")+
  scale_y_continuous(labels = scales::percent)+
  theme_bw()+
  scale_fill_viridis(discrete=T, direction = -1)
dev.off()
```


Step 3: Analyze the effect of monthly air temperature on hypolimnetic DO, VHOD, and temp dynamics
```{r}
many_lake_temp_sum <- with_temp%>%
  pivot_longer(cols = contains("temp", ignore.case = F), names_to = "mon")%>%
  filter(!mon=="mean_temp")%>%
  mutate(mon = month(as.Date(paste0("2022-",substr(mon,6,8),"-01"), format = "%Y-%b-%d")),
         Year = ifelse(mon>=9,Year+1,Year))

# DO
many_lake_stat <- many_lake_temp_sum%>%
  filter(!is.na(DO_mgL_HYPO),!is.na(mon), !is.na(value))%>%
  group_by(LakeID)%>%
  mutate(hypoxic = ifelse(max(DO_mgL_HYPO)<1,"yes","no"))%>%
  filter(hypoxic=="no")%>%
  group_by(LakeID, mon)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
         )%>%
  dplyr::summarize(do_temp = pcor.test(DO_mgL_HYPO,value,Year,method = "spearman")$estimate,
            )

labs = c("N:     Sep\nS:     Mar","Oct\nApr","Nov\nMay","Dec\nJun","Jan\nJul","Feb\nAug","Mar\nSep","Apr\nOct","May\nNov","Jun\nDec","Jul\nJan","Aug\nFeb")

anova_res <- aov(do_temp~as.factor(mon),data=many_lake_stat)
length(unique(many_lake_stat$LakeID))
summary(anova_res)
tukey <- TukeyHSD(anova_res)
cld <- multcompLetters4(anova_res,tukey)$`as.factor(mon)`$Letters
cld_df <- data.frame(cld = cld, mon = names(cld))%>%
  mutate(mon = factor(as.numeric(mon), levels = c(9:12,1:8), labels = labs))


jpeg("../Figures/DO and temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
do_temp_data <- many_lake_stat%>%
  group_by(mon)%>%
  mutate(p = wilcox.test(do_temp)$p.value,
         n = n(),
         mon = factor(mon, levels = c(9:12,1:8),labels = labs))
do_temp <- do_temp_data%>%
  ggplot(aes(x=mon, y = do_temp))+
  geom_boxplot(aes(color=p<0.001))+
  geom_text(aes(x = mon, label = cld, y = 1.08),
            data = cld_df)+
  #geom_text(aes(label = paste0("n = ",n), y = max(do_temp)))+ #n is the same across all
  geom_hline(yintercept=0)+
  ylim(-1,1.08)+
  xlab("Air temp. month")+
  ylab(paste0("Correlation with\nlate-summer DO\n(n = ",unique(do_temp_data$n),")"))+
  theme_bw()+
  scale_color_manual(values = c("grey50","#FB8B24"))+
  theme(axis.text.x = element_text(hjust = ifelse(levels(do_temp_data$mon) == "N:     Sep\nS:     Mar", .8, .5)))
do_temp
dev.off()


######## Hypo temp
# Stratified hypo temp

many_lake_stat <- many_lake_temp_sum%>%
  filter(!is.na(strat_Temp_C_HYPO),!is.na(mon), !is.na(value))%>%
  group_by(LakeID, mon)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
         )%>%
  dplyr::summarize(temp_temp = pcor.test(strat_Temp_C_HYPO,value,Year,method = "spearman")$estimate,
            )
anova_res <- aov(temp_temp~as.factor(mon),data=many_lake_stat)
summary(anova_res)
length(unique(many_lake_stat$LakeID))
tukey <- TukeyHSD(anova_res)
cld <- multcompLetters4(anova_res,tukey)$`as.factor(mon)`$Letters
cld_df <- data.frame(cld = cld, mon = names(cld))%>%
  mutate(mon = factor(as.numeric(mon), levels = c(9:12,1:8), labels = labs))

temp_temp_data <- many_lake_temp_sum%>%
  group_by(LakeID, mon)%>%
  filter(!is.na(strat_Temp_C_HYPO),!is.na(mon), !is.na(value))%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
         )%>%
  dplyr::summarize(temp_temp = pcor.test(strat_Temp_C_HYPO,value,Year,method = "spearman")$estimate,
            )%>%
  group_by(mon)%>%
  mutate(p = wilcox.test(temp_temp)$p.value,
         n = n(),
         mon = factor(mon, levels = c(9:12,1:8), labels = labs))
unique(temp_temp_data$LakeID)

jpeg("../Figures/Air and strat hypo temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
temp_temp <- temp_temp_data%>%
  ggplot(aes(x=mon, y = temp_temp))+
  geom_boxplot(aes(color=p<0.001))+
  geom_text(aes(x = mon, label = cld, y = 1.08),
            data = cld_df)+
  geom_hline(yintercept=0)+
  xlab("Air temp. month")+
  ylab(paste0("Correlation with\nstratified average hypo.\ntemp. (n = ",unique(temp_temp_data$n),")"))+
  theme_bw()+
  ylim(-1,1.08)+
  scale_color_manual(values = c("grey50","#FB8B24"))+
  theme(axis.text.x = element_text(hjust = ifelse(levels(temp_temp_data$mon) == "N:     Sep\nS:     Mar", .8, .5)))
temp_temp
dev.off()

many_lake_stat <- many_lake_temp_sum%>%
  filter(!is.na(DO_demand_mgLd_HYPO),!is.na(mon), !is.na(value))%>%
  group_by(LakeID, mon)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
         )%>%
  dplyr::summarize(temp_demand = pcor.test(DO_demand_mgLd_HYPO,value,Year,method = "spearman")$estimate,
            )
anova_res <- aov(temp_demand~as.factor(mon),data=many_lake_stat)
summary(anova_res)
length(unique(many_lake_stat$LakeID))
tukey <- TukeyHSD(anova_res)
cld <- multcompLetters4(anova_res,tukey)$`as.factor(mon)`$Letters
cld_df <- data.frame(cld = cld, mon = names(cld))%>%
  mutate(mon = factor(as.numeric(mon), levels = c(9:12,1:8), labels = labs))

jpeg("../Figures/Demand and air temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
temp_demand_data <- many_lake_temp_sum%>%
  group_by(LakeID, mon)%>%
  filter(!is.na(DO_demand_mgLd_HYPO),!is.na(mon), !is.na(value))%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
         )%>%
  dplyr::summarize(demand_temp = pcor.test(DO_demand_mgLd_HYPO,value,Year,method = "spearman")$estimate,
            )%>%
  group_by(mon)%>%
  mutate(p = wilcox.test(demand_temp)$p.value,
         n = n(),
         mon = factor(mon, levels = c(9:12,1:8),labels = labs))
   
temp_demand <- temp_demand_data%>%
  ggplot(aes(x=mon, y = demand_temp))+
  geom_boxplot(aes(color=p<0.001))+
  geom_text(aes(x = mon, label = cld, y = 1.08),
            data = cld_df)+
  geom_hline(yintercept=0)+
  xlab("Air temp. month")+
  ylab(paste0("\nCorrelation with\nVHOD (n = ",unique(temp_demand_data$n),")"))+
  theme_bw()+
  ylim(-1,1.08)+
  scale_color_manual(values = c("grey50","#FB8B24"))+
  theme(axis.text.x = element_text(hjust = ifelse(levels(temp_demand_data$mon) == "N:     Sep\nS:     Mar", .8, .5)))
temp_demand
dev.off()

jpeg("../Figures/Monthly correlations.jpg", width = 5, height = 8, res = 300, units = "in")
ggarrange(temp_temp+theme(axis.title.x = element_blank()),temp_demand+theme(axis.title.x = element_blank()),do_temp,
          common.legend = T, legend = "bottom",nrow=3, labels = "auto", align = "hv")
dev.off()
```

