---
title: "Lags"
author: "Abby Lewis"
date: "2023-05-22"
output: html_document
---

This file analyzes relevant lags for ABA relationships of interest. 

Table of contents:
Step 1: Load packages and data
Step 2: Hypo. DO vs. TP lag
Step 3: Hypo. TP vs epi. TP lag
Step 4: Epi. TP vs chl-a lag
Step 5: Chl-a vs VHOD lag
Step 6: VHOD vs hypo. DO lag


Step 1: Load packages and data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

set.seed(47) #Set seed so points jitter the same every time

#Load saved data
with_temp <- read.csv("../Compiled data/All_data_annual.csv")
```


Step 2: Hypo. DO vs. TP lag
```{r}
p_do_stat <- with_temp%>%
  unique()%>%
  group_by(LakeID)%>%
  arrange(LakeID,Year)%>%
  mutate(do_lag = lag(DO_mgL_HYPO),
         do_lag2 = lag(do_lag),
         do_lag2 = ifelse(lag(lag(Year))==Year-2,do_lag2,NA),
         lag_is_last_year = ifelse(lag(Year)==(Year-1),T,F))%>%
  filter(lag_is_last_year)%>%
  pivot_longer(cols = c(DO_mgL_HYPO,do_lag,do_lag2))%>% #, chla_lag3
  filter(!is.na(value),!is.na(TP_ugL_HYPO))%>%
  group_by(name,LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  arrange(LakeID)%>%
  filter(nyear>=10,
         #(max(value)-min(value))>1,
         #min(value)<(0.7*max(value))
         )%>%
  dplyr::summarize(p_do = cor.test(TP_ugL_HYPO,value,method = "spearman")$estimate
    #demand_chla = pcor.test(DO_demand_mgLd_HYPO,value,Year,method = "spearman")$estimate
            )

jpeg("../Figures/Hypo P vs DO lag.jpg", width = 4, height = 3.5, res = 300, units = "in")
p_do_stat%>%
  group_by(name)%>%
  mutate(p = wilcox.test(p_do)$p.value,
         n = n())%>%
  ggplot(aes(x=factor(name, levels = c("do_lag3","do_lag2","do_lag","DO_mgL_HYPO","do_lead","do_lead2"), labels = c("t-3","t-2","t-1","t","t+1", "t+2")),y=p_do))+
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(color = p<0.05), outlier.shape = NA)+
  geom_jitter(width = 0.15, color = "grey20", fill = "white",shape = 21)+
  geom_text(aes(label = paste0("n = ",n), y = max(p_do)+.1))+
  ylab("Correlation between hypo. DO and hypo. TP")+
  scale_color_manual(values = c("grey50","#FB8B24"), breaks = c(F,T),labels = c("FALSE","TRUE"),drop = F, name = "p < 0.05")+
  xlab("Hypo. DO lag")+
  theme_bw()
dev.off()
```


Step 3: Hypo. TP vs epi. TP lag
```{r}
p_p_stat <- with_temp%>%
  unique()%>%
  group_by(LakeID)%>%
  arrange(LakeID,Year)%>%
  mutate(p_lag = lag(TP_ugL_HYPO),
         p_lag2 = lag(p_lag),
         p_lag2 = ifelse(lag(lag(Year))==Year-2,p_lag2,NA),
         lag_is_last_year = ifelse(lag(Year)==(Year-1),T,F))%>%
  filter(lag_is_last_year)%>%
  pivot_longer(cols = c(TP_ugL_HYPO,p_lag,p_lag2))%>% #, chla_lag3
  filter(!is.na(value),!is.na(TP_ugL_EPI))%>%
  group_by(name,LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  arrange(LakeID)%>%
  filter(nyear>=10,
         #(max(value)-min(value))>1,
         #min(value)<(0.7*max(value))
         )%>%
  dplyr::summarize(p_p = cor.test(TP_ugL_EPI,value,method = "spearman")$estimate
    #demand_chla = pcor.test(DO_demand_mgLd_HYPO,value,Year,method = "spearman")$estimate
            )

jpeg("../Figures/Epi P vs hypo P lag.jpg", width = 4, height = 3.5, res = 300, units = "in")
p_p_stat%>%
  mutate(p = wilcox.test(p_p)$p.value,
         n = n())%>%
  ggplot(aes(x=factor(name, levels = c("p_lag3","p_lag2","p_lag","TP_ugL_HYPO","p_lead","p_lead2"), labels = c("t-3","t-2","t-1","t","t+1", "t+2")),y=p_p))+
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(color = p<0.05), outlier.shape = NA)+
  geom_jitter(width = 0.15, color = "grey20", fill = "white",shape = 21)+
  geom_text(aes(label = paste0("n = ",n), y = max(p_p)+.1))+
  ylab("Correlation between hypo. TP and epi. TP")+
  scale_color_manual(values = c("grey50","#FB8B24"))+
  xlab("Hypo. TP lag")+
  theme_bw()
dev.off()
```


Step 4: Epi. TP vs chl-a lag
```{r}
p_chla_stat <- with_temp%>%
  unique()%>%
  group_by(LakeID)%>%
  arrange(LakeID,Year)%>%
  mutate(p_lag = lag(strat_TP_ugL_EPI),
         p_lag2 = lag(p_lag),
         p_lag2 = ifelse(lag(lag(Year))==Year-2,p_lag2,NA),
         lag_is_last_year = ifelse(lag(Year)==(Year-1),T,F))%>%
  filter(lag_is_last_year)%>%
  pivot_longer(cols = c(strat_TP_ugL_EPI,p_lag,p_lag2))%>% #, chla_lag3
  filter(!is.na(value),!is.na(Chla_ugL_EPI))%>%
  group_by(name,LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  arrange(LakeID)%>%
  filter(nyear>=10,
         #(max(value)-min(value))>1,
         #min(value)<(0.7*max(value))
         )%>%
  dplyr::summarize(chla_p = cor.test(Chla_ugL_EPI,value,method = "spearman")$estimate
    #demand_chla = pcor.test(DO_demand_mgLd_HYPO,value,Year,method = "spearman")$estimate
            )

jpeg("../Figures/Epi P vs chl-a lag.jpg", width = 4, height = 3.5, res = 300, units = "in")
p_chla_stat%>%
  mutate(p = wilcox.test(chla_p)$p.value,
         n = n())%>%
  ggplot(aes(x=factor(name, levels = c("p_lag3","p_lag2","p_lag","strat_TP_ugL_EPI","p_lead","p_lead2"), labels = c("t-3","t-2","t-1","t","t+1", "t+2")),y=chla_p))+
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(color = p<0.05), outlier.shape = NA)+
  geom_jitter(width = 0.15, color = "grey20", fill = "white",shape = 21)+
  geom_text(aes(label = paste0("n = ",n), y = max(chla_p)+.1))+
  ylab("Correlation between epi. TP and chl-a")+
  scale_color_manual(values = c("grey50","#FB8B24"), breaks = c(F,T),labels = c("FALSE","TRUE"),drop = F, name = "p < 0.05")+
  xlab("Epi. TP lag")+
  theme_bw()
dev.off()
```


Step 5: Chl-a vs VHOD lag
```{r}
chla_demand_stat <- with_temp%>%
  unique()%>%
  group_by(LakeID)%>%
  arrange(LakeID,Year)%>%
  mutate(do_lag = lag(DO_mgL_HYPO),
         chla_lag = lag(Chla_ugL_EPI),
         chla_lag2 = lag(chla_lag),
         chla_lag2 = ifelse(lag(lag(Year))==Year-2,chla_lag2,NA),
         #chla_lag3 = lag(chla_lag2),
         #chla_lag3 = ifelse(lag(lag(lag(Year)))==Year-3,chla_lag3,NA),
         lag_is_last_year = ifelse(lag(Year)==(Year-1),T,F))%>%
  filter(lag_is_last_year)%>%
  pivot_longer(cols = c(Chla_ugL_EPI,chla_lag, chla_lag2))%>% #, chla_lag3
  filter(!is.na(value),!is.na(DO_demand_mgLd_HYPO))%>%
  group_by(name,LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  arrange(LakeID)%>%
  filter(nyear>=10,
         #(max(value)-min(value))>1,
         #min(value)<(0.7*max(value))
         )%>%
  dplyr::summarize(demand_chla = cor.test(DO_demand_mgLd_HYPO,value,method = "spearman")$estimate
    #demand_chla = pcor.test(DO_demand_mgLd_HYPO,value,Year,method = "spearman")$estimate
            )

jpeg("../Figures/Chla vs Demand lag.jpg", width = 4, height = 3.5, res = 300, units = "in")
chla_demand_stat%>%
  group_by(name)%>%
  mutate(p = wilcox.test(demand_chla)$p.value,
         n = n(),
         color = p<0.05,
         color = factor(color,levels = c(T,F)))%>%
  ggplot(aes(x=factor(name, levels = c("chla_lag3","chla_lag2","chla_lag","Chla_ugL_EPI","chla_lead","chla_lead2"), labels = c("t-3","t-2","t-1","t","t+1", "t+2")),y=demand_chla))+
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(color = color), outlier.shape = NA)+
  geom_jitter(width = 0.15, color = "grey20", fill = "white",shape = 21)+
  geom_text(aes(label = paste0("n = ",n), y = max(demand_chla)+.1))+
  ylab("Correlation between chl-a and VHOD")+
  scale_color_manual(values = c("grey50","#FB8B24"), breaks = c(F,T),labels = c("FALSE","TRUE"),drop = F, name = "p < 0.05")+
  xlab("Chl-a lag")+
  theme_bw()
dev.off()
```


Step 6: VHOD vs hypo. DO lag
```{r}
do_vhod_stat <- with_temp%>%
  unique()%>%
  group_by(LakeID)%>%
  arrange(LakeID,Year)%>%
  mutate(vhod_lag = lag(DO_demand_mgLd_HYPO),
         vhod_lag2 = lag(vhod_lag),
         vhod_lag2 = ifelse(lag(lag(Year))==Year-2,vhod_lag2,NA),
         lag_is_last_year = ifelse(lag(Year)==(Year-1),T,F))%>%
  filter(lag_is_last_year)%>%
  pivot_longer(cols = c(DO_demand_mgLd_HYPO,vhod_lag,vhod_lag2))%>% #, chla_lag3
  filter(!is.na(value),!is.na(DO_mgL_HYPO))%>%
  group_by(name,LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  arrange(LakeID)%>%
  filter(nyear>=10,
         #min(value)<(0.7*max(value))
         )%>%
  dplyr::summarize(do_vhod = cor.test(DO_mgL_HYPO,value,method = "spearman")$estimate
    #demand_chla = pcor.test(DO_demand_mgLd_HYPO,value,Year,method = "spearman")$estimate
            )

jpeg("../Figures/Hypo. DO vs vhod lag.jpg", width = 4, height = 3.5, res = 300, units = "in")
do_vhod_stat%>%
  group_by(name)%>%
  mutate(p = wilcox.test(do_vhod)$p.value,
         n = n(),
         color = p<0.05,
         color = factor(color,levels = c(T,F)))%>%
  ggplot(aes(x=factor(name, levels = c("vhod_lag3","vhod_lag2","vhod_lag","DO_demand_mgLd_HYPO","vhod_lead","vhod_lead2"), labels = c("t-3","t-2","t-1","t","t+1", "t+2")),y=do_vhod))+
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(color = color), outlier.shape = NA)+
  geom_jitter(width = 0.15, color = "grey20", fill = "white",shape = 21)+
  geom_text(aes(label = paste0("n = ",n), y = max(do_vhod)+.1))+
  ylab("Correlation between VHOD and hypo. DO")+
  scale_color_manual(values = c("grey50","#FB8B24"), breaks = c(F,T),labels = c("FALSE","TRUE"),drop = F, name = "p < 0.05")+
  xlab("Hypo. VHOD lag")+
  theme_bw()
dev.off()
```